Multiphysics Core Dynamics Simulation Using the Improved Quasi-Static Method

### 2017-02-18 doc/M&C_2017/FullPaper/MC2017_FullPaper.tex
final version

Other physical quantities, such as temperature, are affected by reaction rates and subsequently affect the operators of the flux equations.  
For IQS, this feedback affects both the shape equation and the parameters of the PRKE; thus, it is an additional nonlinear component to the 
already non-linear shape-amplitude equations.  Each of these components have different temporal behaviors; so it may be beneficial for 
efficiency to evaluate them on different time scales.  The amplitude is more rapidly varying than the shape which is computationally expensive to evaluate
and is evaluated only on macro-time steps. In multiphysics simulations, one may take advantage of a fine-scale power distribution in the coupled
physics components (temperature). Figure~\ref{fig:time} shows a such a solution process for temperature feedback.

@@

The performance of the quasi-statics can also be explained by the computation of the dynamical time scale described by \sct{sect:tau}.  \fig{fig:LRAtc} shows the time scale profile over the transient, computed using \eqt{eq:tau2}.  This plot shows that in a brute force simulation, the flux dominates the time dependent behaivior, while temperature lags in its variance for the majority of the transient.  In an IQS simulation, the time scale behavior of amplitude almost exactly matches the flux, while shape is more varying than temperature throughout most of the transient.  The large $\tau$ for temperature during the beginning of the transient shows that adaptation of the number of updates is important; computational expense on temperature evaluations is being wasted during this time.

@@

The results from \tbl{tab:tran15} show similar performance of IQS with the temperature updates as the LRA.  IQS with 1 temperature update shows a performance that reduces the error to approximately a tenth of the brute force error, and reduces the execution time by about 12\%.  This shows that IQS was able to resolve the nonlinearity between flux and temperature with significantly fewer diffusion evaluations.  Having IQS with 5 updates significantly increased the execution time for the same time step, but the error was reduced.  Comparing this error to a similar brute force error at a smaller time step could show that the runtime was reduced. IQS P-C performed not nearly as well as it did with the LRA benchmark, but still proved to be effective.  Having 5 updates for IQS P-C increased the runtime marginally, but decreased the error significantly.  The transient profile of the varibles' dynamical time scales is shown in \fig{fig:tran15tc}.  This plot exhibits a similar response to that of the LRA. The response of temperature shows that the updates are a computational frugal treatment of the feedback and adaptation of the number of updates is vital for optimization. \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
This work was supported by the Department of Energy, Idaho National Laboratory, and the Integrated University Program Fellowship.  We thank Mark Dehart, Yaqi Wang, the NEAMS program, and INL's Moose/Rattlesnake team for their support.

>>>>

Other physical quantities, such as temperature, are affected by reaction rates and subsequently affect the operators of the flux equations. For IQS, this feedback affects both the shape equation and the parameters of the PRKE; thus, it is an additional nonlinear component to the already non-linear shape-amplitude equations.  Each of these components have different temporal behaviors; so, it may be beneficial for efficiency to evaluate them on different time scales.  The amplitude is more rapidly varying than the shape which is computationally expensive to evaluate and is evaluated only on macro-time steps. In multiphysics simulations, one may take advantage of a fine-scale power distribution in the coupled physics components (temperature). Figure~\ref{fig:time} shows a such a solution process for temperature feedback.

@@

The performance of the quasi-statics can also be explained by the computation of the dynamical time scale described by \sct{sect:tau}.  \fig{fig:LRAtc} shows the time scale profile over the transient, computed using \eqt{eq:tau2}.  This plot shows that in a brute force simulation, the flux dominates the time dependent behavior, while temperature lags in its variance for the majority of the transient.  In an IQS simulation, the time scale behavior of amplitude almost exactly matches the flux, while shape is more varying than temperature throughout most of the transient.  The large $\tau$ for temperature during the beginning of the transient shows that adaptation of the number of updates is important; computational expense on temperature evaluations is being wasted during this time.

@@

The results from \tbl{tab:tran15} show similar performance of IQS with the temperature updates as the LRA.  IQS with 1 temperature update shows a performance that reduces the error to approximately a tenth of the brute force error, and reduces the execution time by about 12\%.  This shows that IQS was able to resolve the nonlinearity between flux and temperature with significantly fewer diffusion evaluations.  Having IQS with 5 updates significantly increased the execution time for the same time step, but the error was reduced.  Comparing this error to a similar brute force error at a smaller time step could show that the runtime was reduced. IQS P-C performed not nearly as well as it did with the LRA benchmark, but still proved to be effective.  Having 5 updates for IQS P-C increased the runtime marginally, but decreased the error significantly.  The transient profile of the variables' dynamical time scales is shown in \fig{fig:tran15tc}.  This plot exhibits a similar response to that of the LRA. The response of temperature shows that the updates are a computational frugal treatment of the feedback and adaptation of the number of updates is vital for optimization. \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
This work was supported by the Department of Energy, Idaho National Laboratory, and the Integrated University Program Fellowship.  We thank Mark Dehart, Yaqi Wang, the NEAMS program, and INL's MOOSE/Rattlesnake team for their support.
